% ---------- ABSTRACT (spec section 4) : page iii, <= 500 words ----------
\phantomsection\addcontentsline{toc}{chapter}{\protect\hspace*{0.5in}Abstract}
\frontheading{Abstract}

Locating people in aerial imagery is central to disaster search and rescue, but
the targets are often only a few pixels wide, partially occluded, and set against
cluttered backgrounds, which defeats standard object detectors. This thesis asks
which architectural additions actually help for this task, and builds a compact
detector around the answer. Starting from a YOLO11m baseline, an additive ablation
introduces one change at a time and evaluates all four configurations on the C2A
disaster-imagery dataset under one training and evaluation protocol. The changes
are channel-and-spatial attention (CBAM) in place of the baseline self-attention
block, a high-resolution P2 detection scale at stride four, and a bidirectional
state-space (Mamba) neck. The results separate the components clearly. The P2 scale is
the effective change: it raises recall of the smallest ($<8$~px) targets from
$0.743$ to $0.757$ and $\mathrm{AP}_{50}$ from $0.843$ to $0.853$ at a cost of about
half a million parameters and one millisecond of latency. CBAM is nearly free and
lowers the parameter count below the baseline while holding accuracy. The
state-space neck adds $2.4$~million parameters and roughly triples inference latency
without improving any accuracy metric, and is therefore excluded from the
recommended model. The recommended CBAM~+~P2 detector, at $19.6$~million parameters,
is competitive with much larger published detectors on the same dataset, matching or
exceeding all but one at about a third of its size, and it runs in real time on a
single consumer GPU. Adapting it to an even smaller backbone for embedded, on-board
use is set out as future work. The main contribution is a controlled, honest account of where
a small-object detection gain comes from for aerial human detection, including a
negative result for the state-space neck, together with a deployable model that
follows from it.
