\chapter{Introduction}

\section{Introduction}
When an earthquake, flood, or fire strikes, the first hours decide how many people
are found alive. Uncrewed aerial vehicles can survey a disaster site far faster
than ground teams, but the imagery they return is hard to search by eye: a person
seen from altitude may occupy only a handful of pixels among rubble, water, or
smoke. Automating the detection of these very small human targets is the problem
this thesis addresses. The work develops a compact detector for people in aerial
disaster imagery, built on the YOLO11 object detector and refined so that it
recovers the smallest targets without growing into a model too large to run on
airborne hardware.

\section{Background / Problem Statement}
General object detection is close to solved for objects that fill a reasonable part
of the frame, but aerial search-and-rescue imagery breaks that assumption. Targets
are tiny, often smaller than eight pixels. In benchmark disaster imagery about
47\% of annotated people occupy fewer than ten pixels~\cite{nihal2024c2a}. They are
partially buried or occluded, scenes are dense, and the background is visually noisy. Standard detectors detect
at strides of eight pixels and coarser, so the smallest people fall through the
grid. Several remedies exist, such as higher-resolution detection scales, feature
attention, and inference-time slicing, but they are usually reported one at a
time and on generic aerial data, and recent state-space (Mamba) modules have been
promoted without a controlled test on this kind of target. It is therefore unclear
which of these additions actually earns its computational cost for very small human
detection. This thesis answers that question directly on a disaster-imagery
benchmark and uses the answer to build a deployable model.

\section{Objectives}
The work is organised around the following objectives.
\begin{itemize}
  \item Build an additive ablation on the YOLO11m baseline that isolates the effect
        of channel-and-spatial attention (CBAM), a high-resolution P2 detection
        scale, and a state-space neck, each under an identical training and
        evaluation protocol.
  \item Identify, from that ablation, the configuration that gives the best recall
        of very small targets at the lowest parameter and latency cost.
  \item Evaluate the resulting model against published detectors on the same
        dataset, and report the state-space neck honestly whether or not it helps.
  \item Deliver a detector compact enough for deployment, and set out how the design
        extends to a smaller backbone for embedded, on-board use.
\end{itemize}

\section{Scope}
The study is built with current open-source tools: the Ultralytics YOLO11
framework, PyTorch, and standard evaluation libraries, all run on a single
consumer GPU. It covers attention, multi-scale detection, and a state-space neck as
architectural changes, and slicing-aided inference as an inference-time setting.
It is limited to the C2A disaster-imagery dataset and to a single object class,
\texttt{person}. Cross-dataset generalisation to other real aerial datasets and the
production of a field-hardened detector are identified as future work.

\section{Unfamiliarity of the Problem}
The problem was approached without reusing a prior solution. The additive ablation,
the choice and placement of components, the state-space neck in the neck rather
than the backbone, and the evaluation protocol were all designed for this work.
Where an existing dataset or module is used, such as C2A, CBAM, or the YOLO11
baseline, it is cited, and the contribution lies in how these are combined,
measured, and reduced to a deployable model.

\section{Project Planning and Work Distribution}
This thesis is the work of a single author, so there is no division of tasks across
team members and a responsibility-assignment matrix does not apply. The work
proceeded as a sequence of phases: benchmarking the YOLO detector family on C2A,
the attention and detection-head ablations, the state-space experiments,
inference-time enhancement, and the lightweight variant, carried out in parallel
with report writing so that writing did not wait on the completion of every
experiment.

\section{Applications of the Work}
The immediate application is aerial search and rescue: a small detector that finds
people in disaster imagery can run on a drone or a laptop at an incident site and
flag likely victims for an operator. The same design applies to other tasks that
depend on spotting very small people from altitude, such as crowd monitoring and
wide-area surveillance, where model size and inference speed matter as much as raw
accuracy.

\section{Organization of the Thesis}
The remainder of the thesis is organised as follows. Chapter~II reviews related
work and defines the base architectures. Chapter~III presents the proposed detector
and the ablation design. Chapter~IV reports the experimental setup, the
quantitative and qualitative results, and the comparison with published work.
Chapter~V discusses the societal, ethical, legal, and environmental context.
Chapter~VI addresses the complex engineering problems and activities involved.
Chapter~VII concludes with a summary, the limitations, and directions for future
work.
