\chapter{Societal, Health, Environment, Safety, Ethical, Legal and Cultural Issues}

\section{Introduction}
A detector intended for disaster response is deployed around vulnerable people
and operated by public agencies, so its wider implications matter as much as
its accuracy. The same model that finds flood victims can, unmodified, watch a
crowd; that dual use runs through every consideration below. This chapter sets
out the intellectual-property, ethical, safety, legal, societal, and
environmental aspects of the work, using the measured resource data recorded
during training rather than generic statements.

\section{Intellectual Property Considerations}
The work builds on external software, data, and published methods, and each is
used within its terms. The detector is built on the Ultralytics YOLO11
framework~\cite{jocher2024yolo11} and PyTorch, both under their open-source
licences; the CBAM design follows its publication~\cite{woo2018cbam}, and the
SAHI library~\cite{akyon2022sahi} handles sliced inference, each credited
where it appears.

The dataset deserves a specific note. C2A is distributed through
Kaggle~\cite{c2akaggle} by its authors, who request citation of the
accompanying paper~\cite{nihal2024c2a} but publish no explicit licence text.
It is therefore used here strictly for academic research, cited in both
forms, and not redistributed. The sources it composites, AIDER
backgrounds~\cite{kyrkou2020aider} and pose imagery segmented with
U$^2$-Net~\cite{qin2020u2net}, carry their own academic-use terms. The
additive ablation design, the trained weights, the negative state-space
finding, and the analysis are the author's contributions; any public release
of weights would state that they derive from C2A and inherit its
academic-use position.

\section{Ethical Considerations}
The system detects people, which obliges care on several fronts. The first is
dual use: a person detector tuned for rescue is technically indistinguishable
from one used for surveillance. This thesis frames, trains, and evaluates the
model for search and rescue, and Section~\ref{sec:legal5} notes the
regulatory line, but the honest position is that misuse cannot be prevented
by architecture, only by governance of deployment.

The second is error asymmetry. A missed victim is a life risk, while a false
alarm costs an operator seconds, so the evaluation weights recall through the
$F_2$ metric and the operating threshold errs towards flagging, with a human
making every final judgement. Two data-side considerations complete the
picture. The training imagery is composited rather than photographic, so no
images of real, identifiable disaster victims were used to develop the model.
At the same time, C2A covers four scene types and composited poses, and the
model may behave differently on populations, clothing, or terrain it never
saw; the cross-dataset validation planned as future work is the concrete
answer to that bias.

\section{Safety Considerations}
The detector is an assistive tool, not an autonomous decision-maker. It
directs an operator's attention and does not clear areas; its outputs must
never be read as a definitive count of survivors, and
Figure~\ref{fig:qualgrid} shows why, since box counts at the review threshold
exceed the annotated count on dense scenes. Safe use also depends on knowing
where the model is weak. The measured failure modes, heavily occluded
targets, few-pixel targets, and merged detections in crowds, are quantified
in Chapter~IV (Table~\ref{tab:persize}, Section~\ref{sec:qual}) so operators
know where trust should thin. The operating point itself is stable: the
$F_2$-optimal confidence sits in a narrow band (0.16 to 0.20) across all
configurations (Table~\ref{tab:thrcal}) and calibration error is near
$0.02$, so the deployed threshold behaves predictably rather than requiring
per-mission tuning.

\section{Legal Considerations}
\label{sec:legal5}
Airborne collection of imagery that contains people is regulated on two
fronts. Drone operation over populated or disaster areas is subject to
civil-aviation permissions on flight, altitude, and no-fly zones; in
Bangladesh these are administered by the Civil Aviation Authority, and
equivalents apply elsewhere. This thesis concerns the detection model rather
than flight operations, but any deployment inherits those rules. Imagery of
identifiable people also engages data-protection regimes such as the EU
General Data Protection Regulation~\cite{gdpr2016}, even when collected in
an emergency; retention, access, and secondary use of rescue footage need
explicit policy before deployment, and the surveillance-capable nature of
the model makes that policy a requirement rather than a courtesy. Within the
research itself, all experimentation used public academic data under the
terms above, and no personal data was collected by the author.

\section{Impact of the Project on Societal, Health, and Cultural Issues}
Faster location of survivors has a direct health benefit: it shortens the time
to reach injured people, which is decisive in the first hours after an
earthquake or flood. The societal value is greatest where trained search teams
are scarce relative to the affected area, which is the situation in many
disaster-prone regions, including Bangladesh, whose flood seasons regularly
displace communities along major rivers. Because the recommended model is
small (19.6~M parameters, Table~\ref{tab:main}) and runs in real time on a
single consumer GPU, it does not presume expensive infrastructure, which
widens the set of agencies that can afford to operate it. Culturally, aerial
imaging of disaster sites must respect the dignity of victims and local norms
around photographing the dead and injured; the human-in-the-loop design keeps
those judgements with people rather than software.

\section{Impact of Project on the Environment and Sustainability}
The environmental cost of this work was measured, not estimated. A tracker
recorded the energy and emissions of every training run, and a two-second
sampler logged GPU power for their full duration; Table~\ref{tab:carbon}
reports the results. The four runs consumed about $10.6$~kWh of GPU energy
and emitted an estimated $9.0$~kg of CO$_2$ under the local grid intensity,
roughly the emissions of a 75~km car journey. Two design choices kept the
figure low: saved weights were re-evaluated instead of retrained whenever a
metric could be recomputed, and the ablation ran one seed per configuration
in this phase. The state-space variant is the environmental outlier as well
as the accuracy one: it produced more than half of the total emissions
($4.65$~kg, $2.6$ times the recommended model's) for no measured gain, which
turns the argument for honest negative results into an environmental one.
Energy-aware reporting of this kind follows the practice argued for in the
literature on the carbon cost of deep learning~\cite{strubell2019energy}, and
the tracking used the CodeCarbon library~\cite{codecarbon}. At inference, the
recommended model's $14.6$~ms latency on a consumer GPU translates to a small
per-image energy cost, and choosing the 19.6~M-parameter design over a model
three times larger (Table~\ref{tab:sota}) reduces the energy drawn by every
future deployment of the system.

\begin{table}[tbp]
  \centering
  \caption[Measured training cost and CO$_2$ of the four configurations.]{Measured training cost of the four configurations: wall-clock time,
  mean GPU board power from the two-second sampler, derived GPU energy, and
  CO$_2$ estimated by the CodeCarbon tracker~\cite{codecarbon} under the
  Bangladesh grid intensity. The state-space variant alone accounts for more
  than half of the total emissions.}
  \label{tab:carbon}
  \begin{tabular}{@{}l r r r r@{}}
    \toprule
    Configuration & Wall clock (h) & Mean power (W) & Energy (kWh) & CO$_2$ (kg) \\
    \midrule
    YOLO11m baseline    & 7.3  & 216 & 1.6 & 1.31 \\
    + CBAM              & 7.0  & 218 & 1.5 & 1.28 \\
    + CBAM + P2         & 10.2 & 207 & 2.1 & 1.78 \\
    + Mamba + CBAM + P2 & 29.5 & 182 & 5.4 & 4.65 \\
    \midrule
    \textbf{Total}      & \textbf{54.0} & --- & \textbf{10.6} & \textbf{9.02} \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Conclusion}
The considerations above do not sit beside the engineering; they shaped it.
The recall-weighted evaluation encodes the ethical asymmetry of the task, the
composited dataset avoids imagery of real victims, the documented failure
modes and stable operating point make the human-in-the-loop role practical,
and the measured carbon account puts a number on what each architectural
decision cost the environment. The dual-use nature of person detection
remains the standing obligation: the model built here to find people in
rubble must be governed, wherever it is deployed, so that it stays on that
side of the line.
